Online {\IterComp}

### 2017-11-08 infra.tex
infrastructure rewrites

Although {\itercomp} had been originally proposed as an \textit{offline} optimization strategy, this work shows that this powerful
technique can be adapted to work in \textit{online} scenarios. Instead of selecting the best optimization sequence at development time, a
first version of the program is shipped bundled with the {\itercomp} mechanism. Our mechanism produces new application binaries using
different optimization sequences. When the user starts the application, we load and execute one of these binaries to evaluate the
corresponding sequence online. Our mechanism operates on the application's Intermediate Representation (IR) produced by LLVM. This saves us
the effort of repeatedly parsing the code through the frontend and removes the need for shipping the full application code.

@@

Figure~\ref{fig:infra-diagram} shows an overview of how our online {\itercomp} mechanism works. On the developer's side, the program is
pre-compiled to IR without optimizations and is instrumented to measure the amount of work performed and the wall-clock time. On the user's
side, we select an optimization sequence and then compile the instrumented IR to binary using this sequence. Next time the user runs the
application, we measure the work and the execution time. We accumulate multiple such measurements, executing each input only once, until we
can estimate the efficiency metric P of this optimization sequence with high confidence. Finally, we use this information to help us select
the next optimization sequence.

This work focuses on the highlighted component of Figure~\ref{fig:infra-diagram}, the \textit{Work Instrumentation}. It is performed once
and offline, right after the IR is generated. Its purpose is to measure the amount of work performed. Conceptually, we use the cost model of
Section~\ref{sec:metric} to estimate a cost for each basic block of the IR and then add instrumentation to increase a global work counter
by that cost every time a basic block is executed. The measured work represents the computational effort of the unoptimized program and it
should depend only on the program and the input, not the applied optimizations. For that reason, we instrument the unoptimized IR before any optimizations modify it. We present a detailed description of work
instrumentation and how to keep it lightweight in Section~\ref{sec:prof}.

Another component, the \textit{Search Driver}, is responsible for guiding the optimization process. Each selected optimization sequence
is used for multiple executions of the program in order to gather enough work efficiency measurements. When we are able to
estimate this work efficiency with a confidence interval narrower than a certain threshold, we can start evaluating a different
optimization sequence. We call the number of executions performed using the same optimization sequence the \textit{Input-Window Size}.

>>>>

@@

    Although {\itercomp} had been originally proposed as an \textit{offline} optimization strategy, this work shows that this powerful
    technique can be adapted to work in \textit{online} scenarios. Figure~\ref{fig:infra-diagram} shows an overview of how our online
    {\itercomp} mechanism works. On the developers' side, we would traditionally produce an application binary optimized for the inputs
    provided by them. In our approach instead, the program is pre-compiled to LLVM Intermediate Representation (IR) without optimizations,
    is instrumented to measure the amount of work performed and the wall-clock time, and is bundled with the {\itercomp} mechanism.
    On the user's side, this mechanism selects an optimization sequence and compiles the instrumented IR to binary using this sequence.
    Next time the user runs the application, we load and execute the latest binary and measure its work and execution time. We accumulate
    multiple such measurements, executing each input only once, until we can estimate the work efficiency of this optimization sequence
    with high confidence. Finally, we use this information to help us select the next optimization sequence.

    In the previous Section, we examined the highlighted component of Figure~\ref{fig:infra-diagram}, the \textit{Work Instrumentation}.
    We perform it once and offline, right after the IR is generated before applying optimizations. Its purpose is to measure the amount of
    work performed. The instrumentation creates a global work counter and adds code to a subset of the basic blocks to increment the counter
    each time a block is executed. We calculate the work contribution of each basic block according to the model of Section~\ref{subsec:workmetric}.
    We then select the subset of instrumented basic blocks and the work increments for each one, as discussed in Sections~\ref{subsec:prof}
    and \ref{subsec:relaxed}.

    Another component, the \textit{Search Driver}, is responsible for guiding the optimization process. Each selected optimization sequence
    is used for multiple executions of the program in order to gather enough work efficiency measurements. When we are able to
    estimate this work efficiency with a confidence interval narrower than a certain threshold, we can start evaluating a different
    optimization sequence. We call the number of executions performed using the same optimization sequence the \textit{Input-Window Size}.
    \FIXME{MORE. How does the new search work?}
